Let's Play Congestion Control

### 2017-09-01 tex/paper.tex
Minor tex changes and sigmoid for std deviation
Realize that too high entropy is the root of all evil...

\usepackage[utf8]{inputenc}
\usepackage{booktabs} % For formal tables
\usepackage{algorithmicx}
\usepackage{todonotes}
\usepackage{svg}

@@

\newcommand\fnurl[2]{%
\href{#2}{#1}\footnote{\url{#2}}%
}

@@

\author{Joachim Fabini} 
\affiliation{%
 \institution{Institute of Telecommunications}
 \streetaddress{Gußhausstraße 25}
 \postcode{1040}
 \city{Vienna} 
 \country{Austria}}
\email{joachim.fabini@tuwien.ac.at}

@@

% The default list of authors is too long for headers}
\renewcommand{\shortauthors}{Bachl et al.}

@@

Let $r_t$ be the reward that was received at time $t$. $V(s_t; \theta_v)$ designates the expected mean reward in state $s_t$ given the parameters of the value network $theta_v$. Then, with $\gamma$ being the roll-off factor, which designates the influence that future reward has on the moving average, (set to $0.99$), we generically define the expected n-step return as 
\begin{align*}
R_t = \left(\left(\sum_{i=0}^{k-1} \gamma^ir_{t+i}\right) + \gamma^k \frac{V(s_{t+k}; \theta_v)}{1-\gamma}\right)\left(1-\gamma\right),
\end{align*}
where $k$ is upper-bounded by $t_\text{max}$. $R$ is simply a moving average of future rewards. However, usual moving averages take into account values from the past while this one uses values from the future; it runs reversely.

The loss function, which the value network tries to minimize, is the square of the difference of the actual mean reward received and the expected mean reward
\begin{align*}
l_{\text{v}_t} = \left(R_t - V(s_t; \theta_v)\right)^2.
\end{align*}

@@

The loss function $l_{v_t}$ of the value network is actually the sum of the square of the difference for each of these four moving averages:
\begin{align*}
l_{\text{v}_t} =& \left(R_{\text{packet}_t} - V_\text{packet}(s_t; \theta_v)\right)^2 \\
&+\left(R_{\text{byte}_t} - V_\text{byte}(s_t; \theta_v)\right)^2 \\
&+\left(R_{\text{delay}_t} - V_\text{delay}(s_t; \theta_v)\right)^2 \\
&+\left(R_{\text{duration}_t} - V_\text{duration}(s_t; \theta_v)\right)^2
\end{align*}

@@

With $H$ being the entropy and $S$ being the skewness, the actor network minimizes the loss
\begin{align*}
l_{\text{a}_t} =& -\log \left( \pi_W\left( a_t-1 < W \leq a_t \given s_t; \theta_a \right)\right)\left(\frac{}{} - \Sigma v_t\right)\\ 
&+ \beta H\left(\text{Lognormal}\left( \mu_\text{normal}, {\sigma_{\text{normal}}}^{2}\right) \right)\\
&- \delta S\left(\text{Lognormal}\left( \mu_\text{normal}, {\sigma_{\text{normal}}}^{2}\right) \right).
\end{align*}

@@

\begin{figure}
\includesvg{figures/Reinforcement_learning_diagram}
\caption{The classic reinforcement learning approach.\protect\footnotemark}
\label{fig:reinforcement}
\end{figure}
\footnotetext{from \url{https://upload.wikimedia.org/wikipedia/commons/1/1b/Reinforcement_learning_diagram.svg}}

Thus we propose the Partial Action Actor Critic Learning (PAL) framework (see \autoref{fig:pal}) in which one action generates a number of partial actions ($leq 1$). Each partial interaction generates a partial reward upon generating with the environment. Upon receiving a partial reward, the agent estimates the current state and triggers a new action. When all partial rewards of one action were received, the agent combines them to form the reward and updates the value and actor networks.

@@

Spam and eggs

>>>>

\usepackage[utf8]{inputenc}
\usepackage{booktabs} % For formal tables
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{algorithm}

\usepackage{todonotes}
\usepackage{svg}

@@

\makeatletter
\newcommand{\StateIndent}[1][3]{%
  \setlength\@tempdima{\algorithmicindent}%
  \Statex\hskip\dimexpr#1\@tempdima\relax}
\algdef{S}[WHILE]{WhileNoDo}[1]{\algorithmicwhile\ #1}%
\makeatother

\newcommand{\algorithmautorefname}{Algorithm}

@@

%\author{Joachim Fabini} 
%\affiliation{%
% \institution{Institute of Telecommunications}
% \streetaddress{Gußhausstraße 25}
% \postcode{1040}
% \city{Vienna} 
% \country{Austria}}
%\email{joachim.fabini@tuwien.ac.at}

@@

% The default list of authors is too long for headers}
\renewcommand{\shortauthors}{Bachl, Zseby}

@@

Let $r_t$ be the reward that was received at time $t$. $V(s_t; \theta_\text{v})$ designates the expected mean reward in state $s_t$ given the parameters of the value network $theta_v$. Then, with $\gamma$ being the roll-off factor, which designates the influence that future reward has on the moving average, (set to $0.99$), we generically define the expected n-step return as 
\begin{align*}
R_t = \left(\left(\sum_{i=0}^{k-1} \gamma^ir_{t+i}\right) + \gamma^k \frac{V(s_{t+k}; \theta_\text{v})}{1-\gamma}\right)\left(1-\gamma\right),
\end{align*}
where $k$ is upper-bounded by $t_\text{max}$. $R$ is simply a moving average of future rewards. However, usual moving averages take into account values from the past while this one uses values from the future; it runs reversely.

The loss function, which the value network tries to minimize, is the square of the difference of the actual mean reward received and the expected mean reward
\begin{align*}
l_{\text{v}_t} = \left(R_t - V(s_t; \theta_\text{v})\right)^2.
\end{align*}

@@

The loss function $l_{v_t}$ of the value network is actually the sum of the square of the difference for each of these four moving averages:
\begin{align*}
l_{\text{v}_t} =& \left(R_{\text{packet}_t} - V_\text{packet}(s_t; \theta_\text{v})\right)^2 \\
&+\left(R_{\text{byte}_t} - V_\text{byte}(s_t; \theta_\text{v})\right)^2 \\
&+\left(R_{\text{delay}_t} - V_\text{delay}(s_t; \theta_\text{v})\right)^2 \\
&+\left(R_{\text{duration}_t} - V_\text{duration}(s_t; \theta_\text{v})\right)^2
\end{align*}

@@

With $H$ being the entropy, the actor network minimizes the loss
\begin{align*}
l_{\text{a}_t} =& -\log \left( \pi_W\left( a_t-1 < W \leq a_t \given s_t; \theta_\text{a} \right)\right)\\
&\left( \left(\frac{R_{\text{byte}_t}}{{R_{\text{duration}_t}}} - \frac{v_{\text{byte}_t}}{{v_{\text{duration}_t}}}\right) - \left( \frac{R_{\text{delay}_t}}{{R_{\text{packet}_t}}}- \frac{v_{\text{delay}_t}}{{v_{\text{packet}_t}}} \right)\right)\\ 
&+ \beta H\left(\text{Lognormal}\left( \mu_\text{normal}, {\sigma_{\text{normal}}}^{2}\right) \right).
\end{align*}

@@

\begin{figure}
\includesvg{figures/Reinforcement_learning_diagram}
\caption{The classic reinforcement learning approach.\protect\footnotemark}
\label{fig:reinforcement}
\end{figure}
\footnotetext{adapted from \url{https://upload.wikimedia.org/wikipedia/commons/1/1b/Reinforcement_learning_diagram.svg}}

Thus we propose the Partial Action Actor Critic Learning (PAL) framework (see \autoref{fig:pal}) in which one action generates a number of partial actions ($leq 1$). Each partial interaction generates a partial reward upon generating with the environment. Upon receiving a partial reward, the agent estimates the current state and triggers a new action. When all partial rewards of one action were received, the agent combines them to form the reward and updates the value and actor networks (see \autoref{alg:pal}).

@@

\begin{algorithm}
\caption{Partial Action Actor Critic Learning -- pseudocode for each learning thread.}
\label{alg:pal}
\begin{algorithmic}
\Loop
	\State $l_\text{actions} \gets  [] $
	\State $l_\text{states} \gets  [] $
	\State $l_\text{values} \gets  [] $
	\State $l_\text{rewards} \gets  [] $
	\State $l_\text{estimatedValues} \gets  [] $
	\State $l_\text{snapshots} \gets  [] $
	\State $t \gets 0$
	\State $s_0 \gets \text{initialState}()$
	\State $\theta_\text{a} \gets \theta_{\text{a}_\text{g}}$
	\State $\theta_\text{v} \gets \theta_{\text{v}_\text{g}}$
	\Repeat
		\State $l_\text{states}.\text{append}(s_t)$
		\State $l_\text{estimatedValues}.\text{append}(V(s_t; \theta_\text{v}))$
		\If{$t \bmod t_\text{max} = 0$}
			\State $\theta_\text{a} \gets \theta_{\text{a}_\text{g}}$
			\State $\theta_\text{v} \gets \theta_{\text{v}_\text{g}}$
			\State $l_\text{snapshots}.\text{append}((\theta_\text{a}, \theta_\text{v}))$
		\EndIf
		\State Sample $a_t$ from the probability distribution
		\State $l_\text{actions}.\text{append}(a_t)$
		\State $l_{\text{partialActions}_t} \gets \text{partialActions}(a_t)$
		\ForAll{partial actions $a_{\text{p}_{nt}}$ in $l_{\text{partialActions}_t}$}
			\State Take partial action $a_{\text{p}_{nt}}$
		\EndFor
		\State Receive partial reward $r_{\text{p}_{nt'}}$
		\StateIndent[3] where $t' \leq t$ and $0 \leq n \leq\#(l_{\text{partialActions}_{t'}})$
		\If{all partial rewards of $a_{t'}$ were received}
			\State $r_{t'} \gets \text{combination of all partial rewards }r_{\text{p}_{nt'}}$
			\State $l_\text{rewards}.\text{append}(r_{t'})$
			\If{$\#\left(l_\text{rewards}\right) > t_\text{max}$ or the episode is over}
				\State $t_\text{end} = \min\left(t_\text{max}, \#\left( l_\text{rewards} \right)\right)$
				\State $\theta_\text{backup} \gets \left(\theta_\text{a}, \theta_\text{v}\right)$
				\State $\theta_\text{a}, \theta_\text{v} \gets l_\text{snapshots}[0]$
				\State $R_i+1 \gets \frac{l_\text{estimatedValues}[t_\text{end}]}{1-\gamma}]$
				\For{$i\gets t_\text{end}-1, 0$}
					\State $R_i \gets r_i + \gamma r_{i+1}$
					\todo[inline]{Calculate the gradients}
					\State Calculate the gradients $d\theta_\text{a}$ and $d\theta_\text{v}$
					\State $\theta_{\text{a}_\text{g}} \gets \theta_{\text{a}_\text{g}} + d\theta_\text{a}$
					\State $\theta_{\text{v}_\text{g}} \gets \theta_{\text{v}_\text{g}} + d\theta_\text{v}$
				\EndFor
				\State $l_\text{actions} \gets l_\text{actions}[t_\text{max}:] $
				\State $l_\text{states} \gets l_\text{states}[t_\text{max}:] $
				\State $l_\text{values} \gets l_\text{values}[t_\text{max}:] $
				\State $l_\text{rewards} \gets l_\text{rewards}[t_\text{max}:] $
				\State $l_\text{estimatedValues} \gets l_\text{estimatedValues}[t_\text{max}:] $
				\State $l_\text{snapshots} \gets l_\text{snapshots}[1:] $
				\State $\theta_\text{a}, \theta_\text{v} = \theta_\text{backup}$
			\EndIf
		\EndIf
		\State Generate $s_{t+1}$ using $r_{\text{p}_{t'}}$
		\State $t \gets t+1$
	\Until{reaching the end the episode}
\EndLoop
\end{algorithmic}
\end{algorithm}
